---
title: "First Year Exam - Bioinformatics"
author: "Eric Jordahl (A59109089)"
format: pdf
---

**Process the Data**

*Read the Data*

First we will read the data frame and output the first 6 lines to check that the data looks correct.

```{r}
c <-read.csv("covid19_variants.csv")
head(c)
```
*Filter out Variants*

Then we will to work with the data by removing variant categories that we do not want, like Other and Total.

```{r}
library(ggplot2)
library(dplyr)
filtered_c <- c %>% filter(variant_name != "Other" & 
                             variant_name != "Total") 
```

*Change Date Format*

We will then change the date column to be a numerical variable rather than a character one 

```{r}
date_num <- as.Date(filtered_c$date)
```

**Plot the Data**

Then we will plot the line graph for the first time with date on the x and percentage on the y.

```{r}
plot1 <- ggplot(data=filtered_c, 
                aes(x=date_num, y=percentage, 
                    col=variant_name)) + 
  geom_line()
plot1
```
*Add Axis Labels*

Now we will add a title and correct the axis labels.

```{r}
plot2 <- plot1 + labs(y="Percentage of Sequence Specimens", 
                      x=" ", 
                      title = "COVID-19 Variants in California")
plot2
```

*Remove Legend Title*

We will then remove the legend title 

```{r}
plot3 <- plot2 + theme(legend.title=element_blank())
plot3
```
*Adjust X Axis Breaks and Labeling*

Now we will adjust the x axis to show each month not years

```{r}
plot4 <- plot3 + scale_x_date(date_breaks="1 month", date_labels="%b-%y") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
plot4
```

*Add data source*

Now we will add a caption with where the data was obtained from

```{r}
plot5 <- plot4 + 
  labs(caption="Data Source:https://data.chhs.ca.gov/") + 
  theme(plot.caption=element_text(hjust=0.5))
plot5
```

